---
title: "Neoantigen analysis"
author: "Tao Wu"
date: "`r Sys.Date()`"
output:   
  rmdformats::readthedown:
    highlight: kate
    lightbox: false
    toc_depth: 3
    mathjax: true
---

```{r neo_analysis-setup, include=FALSE}
options(max.print = "85")
knitr::opts_chunk$set(echo = TRUE, comment = "#>", eval = TRUE, collapse = TRUE,cache = FALSE)
knitr::opts_knit$set(width = 85)
```

Following the observation of widespread drug-induced alternative splicing (AS) events, we next investigated their potential to generate novel immunogenic targets by predicting HLA-binding peptides derived from these differential AS events. Code to process data for jcast and NetMHCpan input can be found in `scripts/processing_for_neo.R`. 

```{r save_neo, eval=FALSE, include=TRUE}
rmats_counts <- readRDS("../data/rmats_filter.rds")
###GSE174498_CEM_L-asparaginase no output
rmats_counts <- rmats_counts %>% filter(sample != "GSE174498_CEM_L-asparaginase")
rmats_counts <- rmats_counts %>% filter((abs(IncLevelDifference) > 0.1) & (FDR <0.05))
all_samples <- unique(rmats_counts$sample)
samples <- readRDS("../data/samples_with_HLA.rds")
all_samples <- intersect(all_samples, samples$unid)

library(doParallel)
library(foreach)
#create the cluster
my.cluster <- parallel::makeCluster(
  10, 
  type = "PSOCK"
)
#register it to be used by %dopar%
doParallel::registerDoParallel(cl = my.cluster)

res <- foreach(
  i = 1:length(all_samples),
  .packages = c("dplyr")
) %dopar% {
  all_files <- list.files(paste0("~/GEO_data/SRA/gse_out/",all_samples[i],
                                 "/jcast_out/binding_out/res/"),pattern = "tsv",
                          full.names = T)
  dt <- sapply(all_files,
               function(x){
                 tmp <- data.table::fread(x,data.table = F)
                 colnames(tmp)[5] <- "score"
                 tmp$binding_type <- gsub(".+//","",x) %>% gsub("_.+","",.)
                 return(tmp)
               },simplify = F)
  dt <- bind_rows(dt) %>% distinct_all()
  dt_filter <- dt %>% 
    select(allele, peptide, percentile, binding_type) %>% 
    filter(percentile < 2) %>% distinct_all()
  dt_filter_summ <- dt_filter %>% 
    group_by(peptide, allele) %>% 
    summarise(both_sb = n()) %>% ungroup() %>% filter(both_sb == 2) ###ba and el < 2
  dt_filter <- dt_filter %>% filter(peptide %in% dt_filter_summ$peptide)
  ###peptide binding
  dt1 <- dt_filter %>% 
    tidyr::pivot_wider(names_from = c(allele, binding_type), values_from = percentile)
  dt_all <- dt %>% 
    select(allele, peptide, percentile, binding_type) %>% 
    filter(peptide %in% dt1$peptide) %>% 
    tidyr::pivot_wider(names_from = c(allele, binding_type), values_from = percentile)
  bc <- apply(dt_all[,2:ncol(dt_all)],1,function(x){sum(x < 2, na.rm = T)})
  dt_all$binding_counts <- bc
  dt_all <- dt_all %>% select(peptide, binding_counts, everything()) %>% 
    arrange(desc(binding_counts))
  
  all_mer <- readRDS(paste0("~/GEO_data/SRA/gse_out/",all_samples[i],
                            "/jcast_out/binding_out/all_mers.rds"))
  ###peptide metadata
  dt2 <- all_mer %>% filter(seq %in% dt_all$peptide)
  dt2 <- dt2 %>% 
    tidyr::separate_wider_delim(seq_id, delim = "|",
                                names = c("sp","UniProt_ID","UniProt_symbol","Gene_ID",
                                          "rMATS_type","rMTS_ID","Chr","Anch_Exon_SE",
                                          "Alt_Exon_SE","strand_phase","Msjc","Tier")) %>% 
    select(-sp) %>% rename(pep_len = type) %>% 
    mutate(Msjc = as.numeric(gsub("r","",Msjc)))
  
  dt3 <- rmats_counts %>% filter(sample %in% all_samples[i])
  dt_meta <- dt2 %>% distinct_all() %>% 
    mutate(rMATS_type2 = sub("[0-9]$", "", rMATS_type)) %>% 
    inner_join(., dt3 %>% rename(rMTS_ID = ID, rMATS_type2 = type) %>% 
                mutate(rMTS_ID = as.character(rMTS_ID))) %>% ##FDR < 0.05 psi > 0.1 
    select(-GeneID) %>% 
    select(seq, pep_len, rMTS_ID, sample, everything())
  dt_all <- dt_all %>% filter(peptide %in% dt_meta$seq)
  saveRDS(dt_meta, 
          paste0("../scripts/Shiny/data/",
                 all_samples[i], "_binding_pep_meta.rds"))
  saveRDS(dt_all, 
          paste0("../scripts/Shiny/data/",
                 all_samples[i], "_binding_pep.rds"))
}
parallel::stopCluster(cl = my.cluster)

####stat
all_files <- list.files("scripts/Shiny/data/",pattern = "_pep.rds")
sample_meta <- readRDS("../data/all_sample_meta.rds")
sample_meta <- sample_meta %>% 
  filter(unid %in% gsub("_binding_pep.rds","",all_files))

library(doParallel)
library(foreach)
#create the cluster
my.cluster <- parallel::makeCluster(
  60, 
  type = "PSOCK"
)
#register it to be used by %dopar%
doParallel::registerDoParallel(cl = my.cluster)

res <- foreach(
  i = sample_meta$unid,
  .packages = c("dplyr")
) %dopar% {
  dt <- readRDS(paste0("../scripts/Shiny/data/",i,"_binding_pep.rds"))
  dt <- dt %>%
    select(-binding_counts) %>% 
    tidyr::pivot_longer(cols = 2:ncol(.), names_to = "allel", values_to = "rank")
  dt_filter <- dt %>% filter(rank < 0.5)
  dt_filter_summ <- dt_filter %>% 
    mutate(allel = gsub("_.+","",allel)) %>% 
    group_by(peptide, allel) %>% 
    summarise(both_sb = n()) %>% ungroup() %>% filter(both_sb == 2) 
  dt_filter <- dt_filter %>% filter(peptide %in% dt_filter_summ$peptide)
  data.frame(
    unid = i,
    sb_counts = length(unique(dt_filter$peptide))
  )
}
parallel::stopCluster(cl = my.cluster)

res <- bind_rows(res)
sample_meta <- left_join(sample_meta, res)

spres <- readRDS("../data/rmats_fdr01.rds")
spres <- spres %>% 
  group_by(ids) %>% summarise(t_counts = sum(counts)) %>% ungroup()
sample_meta <- left_join(sample_meta, spres %>% rename(unid = ids))
saveRDS(sample_meta,"data/sample_neo.rds")
```

Overall, we identified a substantial repertoire of drug-induced HLA-binding peptides across the dataset. The abundance of these peptides varied across samples depending on the tissue origin and the specific drug treatment.

```{r fig.height = 10}
rm(list = ls())
sample_meta <- readRDS("../data/sample_neo.rds")
sample_meta <- sample_meta %>% mutate(log2counts = log2(sb_counts))
ord <- sample_meta %>% group_by(Tissue_Source2) %>% 
  summarise(med = median(log2counts)) %>% 
  ungroup() %>% arrange(desc(med))
p1 <- ggboxplot(sample_meta,x="Tissue_Source2",y="log2counts",add = "jitter",
                color = "Tissue_Source2",order = ord$Tissue_Source2,xlab = F)+
  theme(legend.position = "none")+
  rotate_x_text(45)

dt <- sample_meta %>% filter(!is.na(DrugClass))
ord <- dt %>% group_by(DrugClass) %>% summarise(med = median(log2counts)) %>% 
  ungroup() %>% arrange(desc(med))
p2 <- ggboxplot(dt,x="DrugClass",y="log2counts",add = "jitter",
                color = "DrugClass",order = ord$DrugClass,xlab = F)+
  theme(legend.position = "none")+
  rotate_x_text(90)
p1 / p2
```

To identify the most potent modulators of splicing-derived antigenicity, we ranked the drug treatments by the number of differential AS-derived HLA-binding peptides. Consistent with our previous findings, Indisulam, which induced the most profound global AS alterations, ranked first.

```{r}
dt_summ <- sample_meta %>% group_by(Drug) %>% 
  summarise(counts = length(unique(cell_line))) %>% ungroup() %>% 
  filter(counts >= 3)
dt <- sample_meta %>% filter(Drug %in% dt_summ$Drug)
drug_summ <- dt %>% group_by(Drug) %>% 
  summarise(median_c = median(log2counts)) %>% ungroup() %>% 
  arrange(desc(median_c)) %>% 
  slice_head(n=10)
dt_filter <- dt %>% filter(Drug %in% drug_summ$Drug)

ggboxplot(dt_filter,x="Drug",y="log2counts",add = "jitter",
          color = "Drug",order = drug_summ$Drug, ylab = "Log2Counts")+
  theme(legend.position = "none")+
  rotate_x_text(45)
```

Notably, we observed a strong positive correlation (Pearson R = 0.86) between the number of differential AS events and the number of HLA-binding peptides across drug treatments, indicating that the raw count of binding peptides is largely driven by the overall splicing alteration magnitude. Given this inherent bias, we normalized the peptide counts by the total number of differential AS events per treatment to evaluate the splicing-to-antigen conversion efficiency. Strikingly, after normalization, onvansertib and its combinations emerged as top-ranking drug treatments, despite its relatively modest impact on global AS in previous analyses.

```{r}
sample_meta <- sample_meta %>% mutate(log2AScounts = log2(t_counts))
ggscatter(sample_meta, x = "log2counts", y = "log2AScounts",
          color = "black", shape = 21, size = 3,
          add = "reg.line",  
          add.params = list(color = "blue", fill = "lightgray"),
          conf.int = TRUE, 
          cor.coef = TRUE, 
          cor.coeff.args = list(method = "pearson", label.x = 10, label.sep = "\n"),
          xlab = "Neoantigen Counts", ylab = "Differential AS Counts"
)
```

```{r}
dt <- dt %>% 
  mutate(norm_counts = sb_counts / t_counts)
drug_summ <- dt %>% group_by(Drug) %>% 
  summarise(median_c = median(norm_counts)) %>% ungroup() %>% 
  arrange(desc(median_c)) %>% 
  slice_head(n=10)
dt_filter <- dt %>% filter(Drug %in% drug_summ$Drug)
ggboxplot(dt_filter,x="Drug",y="norm_counts",add = "jitter",
          color = "Drug",order = drug_summ$Drug, 
          ylab = "Normalized Counts (Neoantigen / Differential AS)")+
  theme(legend.position = "none")+
  rotate_x_text(45)

```

This suggests a distinct class of drugs with high splicing-to-antigen conversion rates, the antigenic potential of which is masked when relying solely on raw counts. To rigorously validate this higher conversion efficiency of onvansertib and rule out confounding effects from differing HLA genotypes across cell lines, we restricted our analysis to the shared HLA alleles between indisulam- and onvansertib-treated cell lines. By calculating the ratio of binding peptides to their originating differential AS events for these shared alleles, we found that onvansertib indeed exhibits a substantially higher splicing-to-antigen conversion efficiency compared to indisulam.

```{r save_2drugs, eval=FALSE, include=TRUE}
sample_meta <- readRDS("../data/all_sample_meta.rds")
samples <- readRDS("../data/samples_with_HLA.rds")
onv_sample <- sample_meta %>% filter(Drug == "onvansertib")
onv_hla <- samples %>% filter(unid %in% onv_sample$unid)
indi_sample <- sample_meta %>% filter(Drug == "Indisulam")
indi_hla <- samples %>% filter(unid %in% indi_sample$unid)
overlap_hla <- intersect(paste0(indi_hla$HLA,collapse = ",") %>% strsplit(.,",") %>% `[[`(1),
                         paste0(onv_hla$HLA,collapse = ",") %>% strsplit(.,",") %>% `[[`(1))
all_samples <- c(onv_hla$unid, indi_hla$unid) %>% unique()
rmats_counts <- readRDS("../data/rmats_filter.rds")
rmats_counts <- rmats_counts %>% filter((abs(IncLevelDifference) > 0.1) & (FDR <0.05))
dt_res <- rmats_counts %>% 
  filter(sample %in% all_samples)

res <- vector("list",length(all_samples))
for (i in 1:length(all_samples)){
  dt <- readRDS(paste0("../scripts/Shiny/data/",all_samples[i],"_binding_pep.rds"))
  dt <- dt %>%
    select(-binding_counts) %>% 
    tidyr::pivot_longer(cols = 2:ncol(.), names_to = "allel", values_to = "rank")
  dt_filter <- dt %>% 
    filter(gsub("_.+","",allel) %in% overlap_hla)
  dt_filter <- dt_filter %>% filter(rank < 0.5)
  dt_filter_summ <- dt_filter %>% 
    mutate(allel = gsub("_.+","",allel)) %>% 
    group_by(peptide, allel) %>% 
    summarise(both_sb = n()) %>% ungroup() %>% filter(both_sb == 2) 
  dt_filter <- dt_filter %>% filter(peptide %in% dt_filter_summ$peptide)
  
  dt_meta <- readRDS(paste0("../scripts/Shiny/data/",all_samples[i],
                            "_binding_pep_meta.rds"))
  dt_meta <- dt_meta %>% filter(seq %in% dt_filter$peptide)
  dt_meta <- dt_meta %>% mutate(ids = paste0(rMATS_type2,"_",rMTS_ID))
  
  dt_mer <- readRDS(paste0("~/GEO_data/SRA/gse_out/",all_samples[i],
                           "/jcast_out/binding_out/all_mers.rds"))
  dt_mer <- dt_mer %>% 
    tidyr::separate_wider_delim(seq_id, delim = "|",
                                names = c("sp","UniProt_ID","UniProt_symbol","Gene_ID",
                                          "rMATS_type","rMTS_ID","Chr","Anch_Exon_SE",
                                          "Alt_Exon_SE","strand_phase","Msjc","Tier")) %>% 
    select(-sp) %>% rename(pep_len = type) %>% 
    mutate(rMATS_type2 = sub("[0-9]$", "", rMATS_type))
  dt_mer <- dt_mer %>% mutate(ids = paste0(rMATS_type2,"_",rMTS_ID))
  
  dt3 <- rmats_counts %>% filter(sample %in% all_samples[i]) %>% 
    mutate(ids = paste0(type,"_",ID))
  dt_mer <- dt_mer %>% filter(ids %in% dt3$ids)
  dt_mer <- dt_mer %>% filter(ids %in% dt_meta$ids)
  
  dt_meta$dev_events <- length(unique(dt_mer$ids))
  res[[i]] <- dt_meta
}

res <- bind_rows(res)
saveRDS(res,"data/Indisulam_onvansertib_neoas.rds")
```

```{r}
rm(list = ls())
res <- readRDS("../data/Indisulam_onvansertib_neoas.rds")
res_summ <- res %>% group_by(sample) %>% 
  summarise(neo_counts = length(unique(seq)),
            as_counts = unique(dev_events)) %>% ungroup() %>% 
  mutate(effc = neo_counts/as_counts) %>% 
  mutate(Drug = gsub(".+_","",sample))
ggboxplot(res_summ,x="Drug",y="effc",add = "jitter",color = "Drug",
          ylab = "#(Binding Peptides) / #(Differential AS)")+
  stat_compare_means()+
  scale_color_npg()+
  guides(color = 'none')

```

We then compared the HLA-binding peptide yield across different types of AS events. Our analysis revealed significant differences in the number of binding peptides generated from various differential splicing types.

```{r save_as_neo, eval=FALSE, include=TRUE}
all_files <- list.files("../scripts/Shiny/data/",pattern = "_pep_meta.rds")
sample_meta <- readRDS("../data/all_sample_meta.rds")
sample_meta <- sample_meta %>% 
  filter(unid %in% gsub("_binding_pep_meta.rds","",all_files))

my.cluster <- parallel::makeCluster(
  60, 
  type = "PSOCK"
)
#register it to be used by %dopar%
doParallel::registerDoParallel(cl = my.cluster)

res <- foreach(
  i = sample_meta$unid,
  .packages = c("dplyr")
) %dopar% {
  dt_meta <- readRDS(paste0("../scripts/Shiny/data/",i,"_binding_pep_meta.rds"))
  dt <- readRDS(paste0("../scripts/Shiny/data/",i,"_binding_pep.rds"))
  dt <- dt %>%
    select(-binding_counts) %>% 
    tidyr::pivot_longer(cols = 2:ncol(.), names_to = "allel", values_to = "rank")
  dt_filter <- dt %>% filter(rank < 0.5)
  dt_filter_summ <- dt_filter %>% 
    mutate(allel = gsub("_.+","",allel)) %>% 
    group_by(peptide, allel) %>% 
    summarise(both_sb = n()) %>% ungroup() %>% filter(both_sb == 2) 
  dt_filter <- dt_filter %>% filter(peptide %in% dt_filter_summ$peptide)
  
  dt_meta <- dt_meta %>% filter(seq %in% dt_filter$peptide)
  dt_summ <- dt_meta %>% 
    group_by(rMATS_type2) %>% 
    summarise(counts = length(unique(seq))) %>% ungroup()
  dt_summ$unid <- i
  return(dt_summ)
}
parallel::stopCluster(cl = my.cluster)

res <- bind_rows(res)
saveRDS(res, "../data/AS_type_neo.rds")
```


```{r}
rm(list = ls())
res <- readRDS("../data/AS_type_neo.rds")
res <- res %>% mutate(log2counts = log2(counts))
res_summ <- res %>% group_by(rMATS_type2) %>% 
  summarise(median_c = median(counts)) %>% ungroup() %>% 
  arrange(desc(median_c))
my_c <- list(c("SE","A3SS"),c("SE","RI"),c("SE","MXE"),c("SE","A5SS"),
             c("A3SS","RI"),c("A3SS","MXE"),c("A3SS","A5SS"),
             c("RI","MXE"),c("RI","A5SS"),
             c("MXE","A5SS"))
ggviolin(res,x="rMATS_type2",y="log2counts",add = "boxplot",
          color = "rMATS_type2",order = res_summ$rMATS_type2, 
          xlab = F)+
  theme(legend.position = "none")+
  stat_compare_means(comparisons = my_c)

```

Specifically, SE events yielded the highest number of binding peptides, whereas A5SS events yielded the fewest. Interestingly, while SE contributed the most peptides in total, RI events exhibited the highest proportion of binding peptides (2.07% versus 2.05% for SE). This observation parallels our drug-level findings, reinforcing that the AS event type with the greatest absolute output may differ from the one with the highest antigenic conversion potential.

```{r save_as_neo_per, eval=FALSE, include=TRUE}
rmats_counts <- readRDS("../data/rmats_filter.rds")
rmats_counts <- rmats_counts %>% filter(sample != "GSE174498_CEM_L-asparaginase")
rmats_counts <- rmats_counts %>% filter((abs(IncLevelDifference) > 0.1) & (FDR <0.05))

all_files <- list.files("scripts/Shiny/data/",pattern = "_pep_meta.rds")
sample_meta <- readRDS("../data/all_sample_meta.rds")
sample_meta <- sample_meta %>% 
  filter(unid %in% gsub("_binding_pep_meta.rds","",all_files))

my.cluster <- parallel::makeCluster(
  10, 
  type = "PSOCK"
)
#register it to be used by %dopar%
doParallel::registerDoParallel(cl = my.cluster)

res <- foreach(
  i = sample_meta$unid,
  .packages = c("dplyr")
) %dopar% {
  dt_meta <- readRDS(paste0("../scripts/Shiny/data/",i,"_binding_pep_meta.rds"))
  dt <- readRDS(paste0("../scripts/Shiny/data/",i,"_binding_pep.rds"))
  dt <- dt %>%
    select(-binding_counts) %>% 
    tidyr::pivot_longer(cols = 2:ncol(.), names_to = "allel", values_to = "rank")
  dt_filter <- dt %>% filter(rank < 0.5)
  dt_filter_summ <- dt_filter %>% 
    mutate(allel = gsub("_.+","",allel)) %>% 
    group_by(peptide, allel) %>% 
    summarise(both_sb = n()) %>% ungroup() %>% filter(both_sb == 2) 
  dt_filter <- dt_filter %>% filter(peptide %in% dt_filter_summ$peptide)
  
  dt_meta_filter <- dt_meta %>% filter(seq %in% dt_filter$peptide)
  dt_binding <- dt_meta_filter %>% 
    group_by(rMATS_type2, rMTS_ID) %>% 
    summarise(binding = length(unique(seq))) %>% ungroup()
  
  dt_mer <- readRDS(paste0("~/GEO_data/SRA/gse_out/",i,"/jcast_out/binding_out/all_mers.rds"))
  dt_mer <- dt_mer %>% filter(!(seq %in% dt_meta$seq)) ###ba 或者 el 至少有一个 大于 2
  dt_mer <- dt_mer %>% 
    tidyr::separate_wider_delim(seq_id, delim = "|",
                                names = c("sp","UniProt_ID","UniProt_symbol","Gene_ID",
                                          "rMATS_type","rMTS_ID","Chr","Anch_Exon_SE",
                                          "Alt_Exon_SE","strand_phase","Msjc","Tier")) %>% 
    select(-sp) %>% rename(pep_len = type) %>% 
    mutate(rMATS_type2 = sub("[0-9]$", "", rMATS_type))
  dt3 <- rmats_counts %>% filter(sample %in% i)
  dt_mer <- dt_mer %>% distinct_all() %>% 
    inner_join(., dt3 %>% rename(rMTS_ID = ID, rMATS_type2 = type) %>% 
                 mutate(rMTS_ID = as.character(rMTS_ID))) ##FDR < 0.05 psi > 0.1 
  
  dt_nobind <- dt_mer %>% 
    group_by(rMATS_type2, rMTS_ID) %>% 
    summarise(no_binding = length(unique(seq))) %>% ungroup()
  dt_all <- full_join(dt_binding, dt_nobind)
  dt_all$unid <- i
  return(dt_all)
}
parallel::stopCluster(cl = my.cluster)

res <- bind_rows(res)
saveRDS(res,"data/neo_as_counts.rds")

```

```{r}
rm(list = ls())
res <- readRDS("../data/neo_as_counts.rds")
res <- res %>% 
  tidyr::pivot_longer(cols = c(binding, no_binding), 
                      names_to = "type", values_to = "counts")
res <- res %>% mutate(counts = ifelse(is.na(counts),0,counts))
res_summ <- res %>% 
  group_by(rMATS_type2, type) %>% 
  summarise(total_counts = sum(counts,na.rm = T)) %>% 
  ungroup()
res_summ %>% 
  group_by(rMATS_type2) %>%
  summarise(binding_per = total_counts[type == "binding"]/total_counts[type == "no_binding"])

library(ggalluvial)
ggplot(data = res_summ,
       aes(axis1 = rMATS_type2, axis2 = type,
           y = total_counts)) +
  scale_x_discrete(limits = c("rMATS_type2", "type"), expand = c(.2, .05)) +
  geom_alluvium(aes(fill = type)) +
  geom_stratum() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_minimal()+
  scale_fill_npg()+
  guides(fill = "none")

```

Ultimately, to screen for drugs with the greatest potential to enhance immunotherapy, we restricted our analysis to peptides binding to the three HLA alleles with the highest population frequencies (Sanchez-Mazas et al., 2024), thereby focusing on candidates with broad patient coverage.

```{r save_neo_topHLA, eval=FALSE, include=TRUE}
samples <- readRDS("../data/samples_with_HLA.rds")
top_hlas <- samples %>% 
  filter(grepl("HLA-A*02:01",HLA,fixed = T) | grepl("HLA-A*24:02",HLA,fixed = T) | grepl("HLA-C*04:01",HLA,fixed = T))
all_samples <- unique(top_hlas$unid)

library(doParallel)
library(foreach)
#create the cluster
my.cluster <- parallel::makeCluster(
  60, 
  type = "PSOCK"
)
#register it to be used by %dopar%
doParallel::registerDoParallel(cl = my.cluster)

res <- foreach(
  i = all_samples,
  .packages = c("dplyr")
) %dopar% {
  dt <- readRDS(paste0("../scripts/Shiny/data/",i,"_binding_pep.rds"))
  dt <- dt %>%
    select(-binding_counts) %>% 
    tidyr::pivot_longer(cols = 2:ncol(.), names_to = "allel", values_to = "rank")
  dt_filter <- dt %>% 
    filter(gsub("_.+","",allel) %in% c("HLA-A*02:01","HLA-A*24:02","HLA-C*04:01"))
  dt_filter <- dt_filter %>% filter(rank < 0.5)
  dt_filter_summ <- dt_filter %>% 
    mutate(allel = gsub("_.+","",allel)) %>% 
    group_by(peptide, allel) %>% 
    summarise(both_sb = n()) %>% ungroup() %>% filter(both_sb == 2) 
  dt_filter <- dt_filter %>% filter(peptide %in% dt_filter_summ$peptide)
  
  dt_meta <- readRDS(paste0("../scripts/Shiny/data/",i,"_binding_pep_meta.rds"))
  dt_meta <- dt_meta %>% filter(seq %in% dt_filter$peptide)
  
  gene_exp <- readRDS(paste0("../scripts/Shiny/data/",i,"_deseq.rds"))
  gene_exp <- gene_exp %>% 
    select(symbol, log2FoldChange) %>% 
    group_by(symbol) %>% 
    slice_max(abs(log2FoldChange), with_ties = F, na_rm = T) %>% ungroup() %>% 
    filter(nchar(symbol) > 0) %>% as.data.frame()
  dt_meta <- left_join(dt_meta, gene_exp %>% rename(geneSymbol = symbol))
  dt_meta <- dt_meta %>% group_by(seq) %>% 
    summarise(max_fc = max(log2FoldChange,na.rm = T),
              max_fc_gene = geneSymbol[which.max(log2FoldChange)]) %>% ungroup()
  
  dt_filter <- left_join(dt_filter, dt_meta %>% rename(peptide = seq))
  dt_filter <- dt_filter %>% filter(max_fc > 1)
  tmp_res <- dt_filter %>% 
    mutate(allel = gsub("_.+","",allel)) %>% 
    group_by(allel) %>% summarise(neo_counts = length(unique(peptide)))
  tmp_res$unid <- i
  return(tmp_res)
}
parallel::stopCluster(cl = my.cluster)

res <- bind_rows(res)
saveRDS(res,"../data/topHLA_neo.rds")

```

```{r}
rm(list = ls())
res <- readRDS("../data/topHLA_neo.rds")
sample_meta <- readRDS("../data/all_sample_meta.rds")
res <- left_join(res,sample_meta)
###至少有三个细胞系的药物
res_summ <- res %>% group_by(Drug) %>% summarise(cell_c = length(unique(cell_line))) %>% 
  ungroup() %>% filter(cell_c >= 3)
res_filter <- res %>% filter(Drug %in% res_summ$Drug)
dt <- res_filter %>% 
  group_by(allel, Drug) %>% 
  summarise(mean_neo_c = median(log2((neo_counts)))) %>% ungroup()
dt <- dt %>% 
  tidyr::pivot_wider(names_from = Drug, values_from = mean_neo_c) %>% as.data.frame()
rownames(dt) <- dt$allel
dt$allel <- NULL
dt <- dt %>% t() %>% as.data.frame() %>% 
  arrange(desc(`HLA-A*02:01`)) %>% t() %>% as.data.frame()
library(ComplexHeatmap)
library(circlize)
col_fun = colorRamp2(c(3, 5, 12), c("#FEE5C1", "#FC8C59", "#B70503"))
Heatmap(dt,cluster_rows = F,cluster_columns = F,col = col_fun,
        name = "Median of log2(#Neo)",border="black",
        rect_gp = gpar(col = "white", lwd = 2),column_names_max_height = unit(8, "cm"))

```

