---
title: "transcriptomic rewiring and drug sim"
author: "Tao Wu"
date: "`r Sys.Date()`"
output:   
  rmdformats::readthedown:
    highlight: kate
    lightbox: false
    toc_depth: 3
    mathjax: true
---

```{r trans_and_drug-setup, include=FALSE}
options(max.print = "85")
knitr::opts_chunk$set(echo = TRUE, comment = "#>", eval = TRUE, collapse = TRUE,cache = FALSE)
knitr::opts_knit$set(width = 85)
```

To characterize the global transcriptomic alterations triggered by pharmacological interventions, we systematically analyzed the differential gene expression (DEG) profiles between drug-treated and untreated control cancer cell lines. To elucidate the functional implications of these transcriptomic shifts, we performed Gene Set Enrichment Analysis (GSEA) using 10,490 ontology gene sets from the MSigDB database. The code for DEG and GSEA analysis can be found in `scripts/gene_exp_analysis.R`. Following drug treatments, cell lines exhibited widespread transcriptomic changes, as evidenced by the substantial number of differentially expressed genes (DEGs) was identified. (|log2FC| > 1 and adjusted P < 0.05) (median number of DEGs is 794). The median number of significantly enriched pathways (adjusted P< 0.1) is 923.

```{r}
sample_meta <- readRDS("../data/all_sample_meta.rds")
diff_res <- vector("list",nrow(sample_meta))
for (i in seq_along(diff_res)){
  tmp <- readRDS(paste0("../scripts/Shiny/data/",
                        sample_meta$unid[i],"_deseq.rds"))
  diff_res[[i]] <- data.frame(ids = sample_meta$unid[i], 
                              diff_counts = sum((tmp$padj < 0.05) & (abs(tmp$log2FoldChange) > 1), na.rm = T))
}
diff_res <- bind_rows(diff_res)

gghistogram(diff_res,x="diff_counts",add = "median", rug = TRUE,
            xlab = "Number of DEGs",ylab = "Number of Treatments",fill = "#D9AE2C")+
  annotate("text", x = median(diff_res$diff_counts)+400, y = 160, 
           label = paste0("Median: ",median(diff_res$diff_counts)), 
           size = 5, color = "red")

```

```{r pathway_num, eval=FALSE, include=TRUE}
sample_meta <- readRDS("../data/all_sample_meta.rds")
diff_res <- vector("list",nrow(sample_meta))
for (i in seq_along(diff_res)){
  tmp <- readRDS(paste0("~/GEO_data/SRA/gse_out/",sample_meta$unid[i],"/diff_gsea.rds"))
  diff_res[[i]] <- data.frame(ids = sample_meta$unid[i], 
                              diff_counts = sum((tmp$padj < 0.1), na.rm = T))
}
diff_res <- bind_rows(diff_res)
gghistogram(diff_res,x="diff_counts",add = "median", rug = TRUE,
            xlab = "Number of Enriched Pathways",ylab = "Number of Treatments",
            fill = "#D9AE2C")+
  annotate("text", x = median(diff_res$diff_counts)+300, y = 80, 
           label = paste0("Median: ",median(diff_res$diff_counts)), 
           size = 5, color = "red")
```

Given the ubiquity of these widespread transcriptomic changes, we next sought to evaluate the similarity of drug-induced transcriptomic responses across different treatment conditions. By calculating the Kendall correlation coefficient between the DESeq2 statistic vectors of any two drug-cell line treatments, we found that, overall, the transcriptomic response similarities were notably low across diverse drug-cell line pairs. This indicates that drug-induced transcriptional perturbations are highly context-dependent. When comparing cell lines from the same tissue origin versus those from different tissues, we observed that the drug-induced transcriptomic perturbation similarities were consistently higher within the same tissue type. This suggests that shared tissue-specific transcriptional programs impose a measurable constraint on the cellular response to pharmacological perturbations. Furthermore, the degree of response similarity was strongly influenced by the mechanism of action of the drug classes. For drug categories with diverse and dispersed targets, such as Natural_Product_Derived/TCM and Other_Therapeutic, the transcriptomic response similarities were particularly low.

```{r cal_sim, eval=FALSE, include=TRUE}
###get deseq2 stat
sample_meta <- readRDS("../data/all_sample_meta.rds")
rank_res <- vector("list",nrow(sample_meta))
for (i in seq_along(rank_res)){
  tmp <- readRDS(paste0("~/GEO_data/SRA/gse_out/",sample_meta$unid[i],"/deseq2_diff.rds"))
  tmp <- tmp %>% 
    select(gene_id, stat)
  colnames(tmp)[2] <- paste0("stat_",i)
  rank_res[[i]] <- tmp
}
rank_res <- Reduce(inner_join,rank_res)
rownames(rank_res) <- rank_res$gene_id
rank_res$gene_id <- NULL
saveRDS(rank_res, "../data/deseq2_stat.rds")

##calculate sim
library(doParallel)
library(foreach)
cal_ken_par <- function(var_dt, ncores){
  all_com <- combn(colnames(var_dt), 2, simplify = F)
  all_com <- sapply(all_com, paste, collapse = "-")
  my.cluster <- parallel::makeCluster(
    ncores, 
    type = "PSOCK"
  )
  doParallel::registerDoParallel(cl = my.cluster)
  
  res <- foreach(
    i = all_com,
    .export = c("var_dt"),
    .packages = c("dplyr")
  ) %dopar% {
    com_tmp <- strsplit(i,"-")[[1]]
    cor_tmp <- kendallknight::kendall_cor(var_dt[[com_tmp[1]]],var_dt[[com_tmp[2]]])
    return(data.frame(coms = i, cor = cor_tmp))
  }
  parallel::stopCluster(cl = my.cluster)
  res <- bind_rows(res)
  return(res)
}

cor_res <- cal_ken_par(rank_res,ncores = 60)
saveRDS(cor_res,"../data/deseq2_cor.rds")
```

```{r fig.height = 8}
sample_meta <- readRDS("../data/all_sample_meta.rds")
cor_res <- readRDS("../data/deseq2_cor.rds")
drug_smiles_sim <- readRDS("../data/drug_smiles_sim.rds")
drug_id <- drug_smiles_sim %>% select(id1,Drug) %>% distinct_all()
mapping <- sample_meta %>% 
  select(Drug, DrugClass,Tissue_Source2) %>% 
  mutate(ids = paste0("stat_",1:nrow(sample_meta))) %>% 
  rename(Tissue = 3) %>% 
  left_join(.,drug_id %>% rename(drug_id = id1)) %>% select(-Drug)

cor_res <- cor_res %>% 
  tidyr::separate_wider_delim(cols = coms, delim = "-", names = c("id1","id2"))
cor_res <- left_join(
  cor_res, mapping %>% 
    rename(id1 = ids, Drug_class1 = DrugClass, Tissue1 = Tissue, Drug1 = drug_id)
) %>% left_join(., mapping %>% 
                  rename(id2 = ids, Drug_class2 = DrugClass, 
                         Tissue2 = Tissue, Drug2 = drug_id))
###NA 是未知药物类型
dt <- cor_res %>% 
  filter(!is.na(Drug_class1)) %>% filter(!is.na(Drug_class2)) %>% 
  mutate(is_same_drug = ifelse(Drug_class1 == Drug_class2,"Y","N")) %>% 
  mutate(is_same_tissue = ifelse(Tissue1 == Tissue2,"Y","N"))
###相同类型的药物在同一组织来源的细胞系和不同组织来源的细胞系之间的相似性
dt1 <- dt %>% filter(is_same_drug == "Y")
dt1 <- dt %>% filter(Drug1 != Drug2)
ggviolin(dt1, x="is_same_tissue",y="cor",facet.by = "Drug_class1", 
         add = "boxplot",color = "is_same_tissue",
         palette = c("#80A9C9", "#89C7C1"),
         add.params = list(outliers = FALSE),
         xlab = "Same Type of Tissue",ylab = "Similarity",legend = "none")+
  stat_compare_means(label = "p.format")+
  theme(
    strip.text = element_text(size = 6, face = "bold"),
    strip.background = element_rect(fill = "lightblue")
  )

```

We also check the distribution of similarity score between drug treatments in the same type of tissue. We found that drug classes with highly specific and similar targets, such as Kinase_inhibitors, exhibited notably higher transcriptomic response similarities.

```{r}
dt <- cor_res %>% 
  filter(Drug_class1 == Drug_class2)
dt1 <- dt %>% filter(Tissue1 == Tissue2) %>% filter(Drug1 != Drug2)
dt2 <- dt %>% filter(Tissue1 != Tissue2)

dt1_summ <- dt1 %>% group_by(Drug_class1) %>% 
  summarise(median_cor = median(cor)) %>% ungroup() %>% 
  arrange(median_cor)
dt1$Drug_class1 <- factor(dt1$Drug_class1, levels = dt1_summ$Drug_class1)
ggplot(dt1, aes(Drug_class1, cor, fill = Drug_class1)) +
  geom_rain(alpha = .5, point.args = list(alpha = 0),
            boxplot.args.pos = list(
              width = 0.1, position = position_nudge(x = 0.13)),
            violin.args.pos = list(
              side = "r",
              width = 0.7, position = position_nudge(x = 0.2))) +
  theme_pubr() +
  scale_fill_npg() +
  guides(fill = 'none', color = 'none') +
  coord_flip()+
  labs(title = 'Drug Treated on Same Type of Tissue',
       y = "Similarity",x="Drug Class")

```

Similarly, samples treated with the same class of drug are enriched in groups with high transcriptional response similarity (Kendall correlation > 0.5).

```{r}
dt <- cor_res %>% filter(Tissue1 == Tissue2)
dt <- dt %>% 
  filter(!is.na(Drug_class1)) %>% filter(!is.na(Drug_class2)) %>% 
  mutate(is_same_drug = ifelse(Drug_class1 == Drug_class2,"Same Drug Class",
                               "Not Same Drug Class")) %>% 
  mutate(cor_type = case_when(
    cor >= 0.5 ~ "High \nTranscriptome Response Similarity",
    cor < 0.5 ~ "Low \nTranscriptome Response Similarity"
  ))
###fisher检验
dt_summ <- dt %>%
  group_by(cor_type,is_same_drug) %>% summarise(counts=n()) %>% ungroup()

ggplot(data=dt_summ,aes(x=cor_type,y=counts,fill=is_same_drug))+
  geom_bar(stat = "identity",position="fill")+
  theme_pubr()+
  labs(y="Percent of cases (%)",title = "Chi-squared test, P-value < 2.2e-16")+
  scale_fill_manual(values=c("#3B4992FF","#EE0000FF"))+
  theme(axis.title.x = element_blank())+
  theme(legend.title = element_blank())

```

While the aforementioned drug classifications are predominantly based on mechanisms of action or molecular targets, current phenotypic drug discovery (Xing et al., 2026) increasingly relies on the inherent structural characteristics of drug molecules themselves. Therefore, to directly assess how the chemical properties of drugs relate to their transcriptomic perturbations, we compared the molecular similarity of drugs between the high and low transcriptomic response similarity groups. We utilized two metrics: the classic Tanimoto similarity based on SMILES representations, and the cosine similarity of molecular embeddings derived from MoLFormer (Ross et al., 2022), a large-scale chemical language model (code for get MolFormer embedding can be found in `scripts/MolFormer_embedding.ipynb`).

```{r cal_drug_sim, eval=FALSE, include=TRUE}
library(rcdk)
library(fingerprint)
drug_meta <- readxl::read_xlsx("data/drug_meta.xlsx")
drug_meta <- drug_meta %>% filter(!is.na(SMILE)) %>% select(-DrugType) %>% distinct_all()
drug_meta <- drug_meta %>% mutate(id = paste0("Drug_",1:nrow(.)))
mols <- parse.smiles(drug_meta$SMILE)
names(mols) <- drug_meta$id
mols <- mols[lengths(mols) != 0]
mols_can <- lapply(mols, get.smiles, flavor = smiles.flavors("Canonical")) %>% unlist()
mols <- sapply(mols_can, parse.smiles)
names(mols) <- names(mols_can)
fp <- lapply(mols, get.fingerprint, type='standard')
fp.sim <- fingerprint::fp.sim.matrix(fp, method='tanimoto')
rownames(fp.sim) <- names(mols)
colnames(fp.sim) <- names(mols)

fp.sim <- as.data.frame(fp.sim)
fp.sim$id1 <- rownames(fp.sim)
fp.sim <- fp.sim %>% tidyr::pivot_longer(cols = 1:248, names_to = "id2", values_to = "sim")
fp.sim <- left_join(fp.sim, drug_meta %>% select(Drug, id) %>% rename(id1 = id))
saveRDS(fp.sim, "data/drug_smiles_sim.rds")

####molformer
dt <- mols_can %>% as.data.frame() %>% rename(smiles = 1)
dt$id <- rownames(dt)
write.csv(dt,"data/drug_smiles.csv",quote = F,row.names = F)
####see MolFormer_embedding.ipynb
molformer_sim <- read.csv("data/drug_molformer_sim.csv")
colnames(molformer_sim)[1] <- "id1"
molformer_sim <- molformer_sim %>% 
  tidyr::pivot_longer(cols = 2:249, names_to = "id2",values_to = "sim")
molformer_sim <- left_join(molformer_sim, 
                           drug_meta %>% select(Drug, id) %>% rename(id1 = id))
saveRDS(molformer_sim, "data/drug_molformer_sim.rds")
```

Both metrics consistently demonstrated that drug pairs within the high transcriptomic similarity group possessed significantly higher chemical structural similarity compared to those in the low-similarity group.

```{r}
cor_res_smile <- left_join(cor_res,
                           drug_smiles_sim %>% 
                             rename(Drug1 = id1, Drug2 = id2) %>% 
                             select(Drug1, Drug2, sim))
cor_res_smile <- cor_res_smile %>% filter(!is.na(sim))

drug_molformer_sim <- readRDS("../data/drug_molformer_sim.rds")
cor_res_molformer <- left_join(cor_res,
                               drug_molformer_sim %>%
                                 rename(Drug1 = id1, Drug2 = id2) %>%
                                 select(Drug1, Drug2, sim))
cor_res_molformer <- cor_res_molformer %>% filter(!is.na(sim))

dt1 <- cor_res_smile %>% 
  filter(Tissue1 == Tissue2) %>% 
  mutate(cor_type = case_when(
    cor >= 0.5 ~ "High",
    cor < 0.5 ~ "Low"
    ))
p1 <- ggboxplot(dt1, x="cor_type",y="sim",
                xlab = "Transcriptome Response Similarity",
                ylab = "SMILEs Similarity (Tanimoto)",
                fill = "cor_type",palette = c("#D9AE2C","#D88C27"))+
  theme(legend.position = "none")+
  stat_compare_means()

dt2 <- cor_res_molformer %>% 
  filter(Tissue1 == Tissue2) %>% 
  mutate(cor_type = case_when(
    cor >= 0.5 ~ "High",
    cor < 0.5 ~ "Low"
  ))
p2 <- ggboxplot(dt2, x="cor_type",y="sim",
                xlab = "Transcriptome Response Similarity",
                ylab = "SMILEs Embedding Similarity (MoLFormer)",
                fill = "cor_type",palette = c("#D9AE2C","#D88C27"))+
  theme(legend.position = "none")+
  stat_compare_means()

p1+p2
```

To quantitatively evaluate the predictive power of structural similarity for transcriptomic convergence, we constructed Receiver Operating Characteristic (ROC) curves using the two structural metrics as predictors. Both features demonstrated robust predictive capabilities, with the MoLFormer embedding similarity achieving a higher AUC (0.814) compared to the Tanimoto similarity (0.791). This finding indicates that structurally analogous compounds, especially those sharing deep chemical language representations, are more likely to induce convergent transcriptomic reprogramming.

```{r}
dt1 <- dt1 %>% mutate(truth = factor(cor_type, levels = c("High","Low")))
dt2 <- dt2 %>% mutate(truth = factor(cor_type, levels = c("High","Low")))

p1 <- yardstick::roc_curve(dt1, truth, sim) %>% 
  ggplot(aes(x = 1 - specificity, y = sensitivity)) +
  geom_path() +
  geom_abline(lty = 3) +
  coord_equal() +
  theme_pubr()+
  annotate("text", x = 0.25, y = 0.9, 
           label = paste0("AUROC: ",round(yardstick::roc_auc(dt1, truth, sim)$.estimate,3)), 
           size = 5, color = "red")
p2 <- yardstick::roc_curve(dt2, truth, sim) %>% 
  ggplot(aes(x = 1 - specificity, y = sensitivity)) +
  geom_path() +
  geom_abline(lty = 3) +
  coord_equal() +
  theme_pubr()+
  annotate("text", x = 0.25, y = 0.9, 
           label = paste0("AUROC: ",round(yardstick::roc_auc(dt2, truth, sim)$.estimate,3)), 
           size = 5, color = "red")
p1 + p2
```

To demonstrate the utility of our database, we sought to identify drugs capable of transforming the tumor immune microenvironment (TME) into a "hot" state, which is generally associated with enhanced anti-tumor immunity and improved responsiveness to immunotherapy. We established a screening strategy based on three gene signatures reflecting distinct immune-modulatory features: antigen processing and presentation machinery from msigdb database, a CRISPR screen-derived gene set of cancer-intrinsic evasion of T cell killing (Lawson et al., 2020) , and genes up-regulated in immune-resistant cancer cell line (Lin et al., 2007). We hypothesized that an ideal immunostimulatory drug would positively enrich the first two gene sets while negatively enriching the immune evasion upregulated gene set. Applying GSEA analysis on our DRIVE data, we evaluated the drug-induced enrichment profiles across various treated cell lines. As a result, we identified multiple drugs that met these criteria.

```{r fig.height = 15}
rm(list = ls())
immune_res <- readRDS("../data/immune_path_gsea.rds")
sample_meta <- readRDS("../data/all_sample_meta.rds")

plot_nes <- function(pathway_gsea, sample_meta_dt, pathway_name, need_drugs){
  pathway_gsea <- left_join(sample_meta_dt, pathway_gsea %>% rename(unid = id))
  dt_sig <- pathway_gsea %>%
    group_by(Drug) %>% mutate(treat_counts = length(unique(cell_line))) %>% 
    ungroup() %>% filter(treat_counts > 3) %>% 
    filter(Drug %in% need_drugs)
  #####
  all_drugs <- unique(dt_sig$Drug)
  drug_nes <- sapply(all_drugs,
                     function(x){
                       tmp <- dt_sig %>% filter(Drug == x)
                       tmp_dt <- data.frame(Drug = x, NES = tmp$NES) %>% arrange(desc(NES))
                       pc <- sum(tmp_dt$NES > 0)
                       nc <- sum(tmp_dt$NES < 0)
                       tmp_dt$y <- c(seq(from=0.5,by=1,length.out=pc),
                                     seq(from=-0.5,by=-1,length.out=nc))
                       return(tmp_dt)
                     },simplify = F)
  drug_nes <- bind_rows(drug_nes)
  
  ggplot(drug_nes, aes(x = Drug, y = y, fill = NES)) + 
    geom_tile(width = 0.5, height = 1, color = "black", size = 0.5)+
    scale_fill_gradient(low = "blue", high = "red") +  
    labs(x = "Drug",
         y = "Cells",
         fill = "NES",
         title = pathway_name)+
    theme(
      axis.text.y = element_blank(),    
      axis.ticks.y = element_blank(),   
      axis.line.y = element_blank()     
    )+
    geom_hline(yintercept = 0,  
               color = "black",    
               linewidth = 1,  
               linetype = "solid")+
    theme(
      panel.background = element_blank(),  
      plot.background = element_blank(),    
      panel.grid = element_blank()        
    )+
    theme(
      axis.text.x = element_text(
        size = 14,         
        angle = 90,         
        hjust = 1,           
        #vjust = 1          
      )
    )
}

dt <- left_join(sample_meta, immune_res %>% rename(unid = id))
dt_summ <- dt %>% 
  group_by(Drug,pathway) %>% 
  summarise(pc = sum(NES > 0), nc = sum(NES < 0)) %>% ungroup() %>% 
  rowwise() %>% 
  mutate(tc = pc+nc) %>% ungroup() %>% filter(tc > 3) %>% 
  rowwise() %>% 
  mutate(is_need = case_when(
    (pathway == "DOWN_Immune_Escape") & (pc > nc) ~ "yes",
    (pathway == "UP_Immune_Escape") & (pc < nc) ~ "yes",
    (pathway == "KEGG_ANTIGEN_PROCESSING_AND_PRESENTATION") & (pc > nc) ~ "yes",
    TRUE ~ "no"
  )) %>% ungroup() %>% 
  group_by(Drug) %>% summarise(yes_c = sum(is_need == "yes")) %>% ungroup() %>% 
  filter(yes_c == 3) %>% 
  filter(!(Drug %in% c("A","T")))

dt <- immune_res %>% 
  filter(pathway == "KEGG_ANTIGEN_PROCESSING_AND_PRESENTATION")
p1 <- plot_nes(dt, sample_meta, "APM", dt_summ$Drug)
dt <- immune_res %>% 
  filter(pathway == "DOWN_Immune_Escape")
p2 <- plot_nes(dt, sample_meta, "DOWN_Immune_Escape", dt_summ$Drug)
dt <- immune_res %>% 
  filter(pathway == "UP_Immune_Escape")
p3 <- plot_nes(dt, sample_meta, "UP_Immune_Escape", dt_summ$Drug)
p1 / p2 / p3

```

Among the identified candidates, We selected Osimertinib for further validation, due to the scarcity of matched pre- and post-treatment data for the other drugs in public repositories. To validate its potential to activate the TME, we leveraged publicly available single-cell RNA sequencing (scRNA-seq) data from (Maynard et al., 2020), which included samples before drug treatment (n=15) and after Osimertinib treatment (n=10). Specifically, the proportion of T cells significantly increased in the post-treatment samples compared to the naive baseline, suggesting an enhanced immune infiltration.

```{r fig.height = 8}
library(Seurat)
rm(list = ls())
sample_meta <- readxl::read_xlsx("../data/CELL_metadata.xlsx",sheet = 2)
tn <- sample_meta %>% filter(`Treatement Timepoint` == "TN")
osi <- sample_meta %>% filter(Treatment == "osimertinib") %>% 
  filter(`Treatement Timepoint` != "TN")

load("../data/IM01_Immune_Seurat_object_nodups-006.RData")
new_obj <- UpdateSeuratObject(tiss_immune)
sub_obj <- subset(new_obj,subset = sample_name %in% c(tn$`Sample Name`,osi$`Sample Name`))

sub_obj$sample_type <- case_when(
  sub_obj$sample_name %in% tn$`Sample Name` ~ "Naive",
  sub_obj$sample_name %in% osi$`Sample Name` ~ "Osimertinib Treatment"
)
sub_obj <- RunUMAP(sub_obj, dims = 1:50, reduction = "pca", reduction.name = "umap")
DimPlot(sub_obj, reduction = "umap", 
        group.by = c("sample_type","immune_subtype_annotation"),ncol=1)

obj_meta <- sub_obj@meta.data
obj_meta <- obj_meta %>% 
  select(sample_name, cell_id, immune_subtype_annotation, sample_type)
obj_summ <- obj_meta %>% 
  group_by(sample_type,sample_name) %>% 
  summarise(t_p = mean(immune_subtype_annotation == "T-cells")) %>% ungroup()
ggboxplot(obj_summ,x="sample_type",y="t_p",ylab = "Percentage of T cells",
          add = "jitter", fill = "sample_type",palette = c("#707DA6","#CCAD9D"),
          xlab = F) +
  stat_compare_means()+theme(legend.position = "none")

unloadNamespace("Seurat")
```

Furthermore, using bulk RNA-seq data of paired patient samples from (Roper et al., 2020), we assessed the functional state of the immune microenvironment. Although the limited sample size restricted statistical power, the GSVA score for the "T CELL MEDIATED CYTOTOXICITY" gene set exhibited a clear increasing trend following Osimertinib treatment. 

```{r}
rm(list = ls())
exp <- readxl::read_xlsx("../data/cell_reports_med_exp.xlsx") %>% 
  as.data.frame()
meta <- readRDS("../data/cell_rep_med_meta.rds")
meta <- meta %>% 
  tidyr::pivot_longer(cols = c(pre,post), names_to = "type", values_to = "sample") %>% 
  mutate(sample_id = paste0(sample,"R"))
meta <- meta %>% distinct_all()

rownames(exp) <- exp$Gene
exp$Gene <- NULL
exp_TPM <- exp %>% mutate(across(everything(), ~(./sum(.,na.rm = T))*10**6))
exp_TPM <- log2(as.matrix(exp_TPM) + 1) %>% as.data.frame()

cyt <- fgsea::gmtPathways("../data/GOBP_T_CELL_MEDIATED_CYTOTOXICITY.v2026.1.Hs.gmt")
para <- GSVA::gsvaParam(exp_TPM %>% as.matrix(), cyt,
                        minSize=5, maxSize=500, kcdf="Gaussian")
res <- GSVA::gsva(para)
res <- t(res) %>% as.data.frame()
res$sample <- rownames(res)
colnames(res)[1] <- "score"

res <- left_join(res %>% rename(sample_id = sample), meta)
dt <- res %>% dplyr::filter(!is.na(sample))
dt_summ <- dt %>% group_by(patient) %>% summarise(counts = length(unique(type))) %>% 
  ungroup() %>% filter(counts == 2)
dt <- dt %>% filter(patient %in% dt_summ$patient)

dt <- dt %>%
  select(patient, type, score) %>% 
  tidyr::pivot_wider(names_from = type, values_from = score)
ggpaired(dt, cond1 = "pre", cond2 = "post",
         fill = "condition", palette = "jco", 
         ylab = "GSVA score of T_CELL_MEDIATED_CYTOTOXICITY")+
  stat_compare_means(paired = T)+
  scale_fill_npg()+
  guides(fill = 'none', color = 'none')
```

Consistent with this, the TIDE score (Jiang et al., 2018), a metric indicative of immune evasion, showed a noticeable decreasing trend in post-treatment. Despite the lack of formal statistical significance, these consistent directional shifts across independent metrics further support the notion that Osimertinib could potentially promote a hotter TME, underscoring the reliability of our database for drug repurposing and immunological research.

```{r}
###TIDE res
tide <- data.table::fread("../data/cell_rep_med_res.txt",data.table = F)
res <- left_join(tide %>% rename(sample_id = V1), meta)
res <- res %>% dplyr::filter(!is.na(sample))
res_summ <- res %>% group_by(patient) %>% summarise(counts = length(unique(type))) %>% 
  ungroup() %>% filter(counts == 2)
res <- res %>% filter(patient %in% res_summ$patient)

dt <- res %>%
  select(patient, type, TIDE) %>% 
  rename(score = 3) %>% 
  tidyr::pivot_wider(names_from = type, values_from = score)
ggpaired(dt, cond1 = "pre", cond2 = "post",
         fill = "condition", ylab = "TIDE score")+
  stat_compare_means(paired = T)+
  scale_fill_npg()+
  guides(fill = 'none', color = 'none')
```



